---
title: "Data Wrangling with dplyr"
author: "Peter Yang"
date: "1/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subset rows and filter columns

```{r load_packages, message=FALSE}

library(dplyr)
library(tidyverse)

```

```{r load_data}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim = "\t")
```

Goals:

subset rows of a dataframe and select columns of a dataframe

```{r}
research <- filter(counts_raw, articleType == "Research Article")
```

Filter to research articles from 2006

```{r}
research_2006 <- filter(research, year == 2006)
nrow(research_2006)
```

Filter to articles with a tweet

```{r}
research_2006_tweet <- filter(research_2006, backtweetsCount  > 0)
nrow(research_2006_tweet)
```

Research articles from 2006 with a FB comment
"|" is "or" sign

```{r}
research_2006_fb_tweet_disease <- filter (research, year == 2006, facebookCommentCount > 0 | backtweetsCount > 0, grepl("Infectious Diseases", plosSubjectTags))
```

Subset the columns with the article information using the dplyr function select()

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
colnames(article_info)
```

New dataframe with columns doi:authorsCount

```{r}
article_info <- select(research, doi:authorsCount)
colnames(article_info)
```

Create a metrics dataframe 

```{r}
metrics <- select(research, contains("Count"))
colnames(metrics)
```

Exclude a column after contains

```{r}
metrics <- select(research, contains("Count"), -authorsCount)
colnames(metrics)
```

```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

```{r}
head(select(research, 3))
```

Slice: slicing only the rows we want by numbers

Rows 1:3 article info dataframe

```{r}
slice(article_info, 1:3)
```

```{r}
low_cite <- filter(research, year <= 2008, pdfDownloadsCount > 1000, mendeleyReadersCount > 15, wosCountThru2011 < 10)
nrow(low_cite)
```

```{r}
nrow(filter(research, year == 2009))
nrow(filter(research, year == 2009, backtweetsCount > 0 | facebookCommentCount > 0))
#nrow(filter(research, year == 2009, facebookCommentCount > 1))
```

```{r}
str(low_cite)
```

# Chaining commands with dplyr

Remember in unix we pipe with the | sign. In dplyr we use %>%
Use shift + control + M 

Facebook articles from 2006

```{r}
facebook_2006 <- research %>% filter(year == 2006) %>% select(contains("facebook"))
head(facebook_2006)
```

```{r}
research_2006 <- filter(research, year == 2006)
facebook_2006 <- select(research_2006, contains("facebook"))
```

Pipe with base R functions as well

```{r}
research %>% filter(year == 2006) %>% select(contains("facebook")) %>% head
```

Arrange functions: use to sort rows in a dataframe

```{r}
research %>% arrange(authorsCount, wosCountThru2011) %>% select(authorsCount, wosCountThru2011) %>% slice(1:10)
```


Flip arrangement of the sort

```{r}
### research %>% arrange(desc(authorsC...))
```

```{r}
research %>% arrange(desc(wosCountThru2011)) %>% select(title, wosCountThru2011) %>% slice(1:3)
research %>% arrange(desc(authorsCount)) %>% select(title, journal, plosSubjectTags, authorsCount) %>% slice(1:3)
```

# Summarize data with dplyr

* new columns with mutate
* sumarize function
* count number of observations m()
* group data by variable (group_by)

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7)

research %>% select(daysSincePublished) %>% head()

```

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7, yearsSincePublished = weeksSincePublished / 52)

select(research, contains("Since")) %>% slice(1:5)
```

Summarize the data

```{r}

summarize(research, plos_mean = mean(plosCommentCount))
```

```{r}
plot_stats <- research %>% summarise(plos_mean = mean(plosCommentCount), 
                                     plos_sd = sd(plosCommentCount))

head(plot_stats)
```

```{r}
research %>% filter(journal == "pone", year == 2007) %>% summarise(plos_mean = mean(plosCommentCount), plos_sd = sd(plosCommentCount))
```

Add number of observations

```{r}
research %>% filter(journal == "pone", year == 2007) %>% summarise(plos_mean = mean(plosCommentCount), 
                                                                   plos_sd = sd(plosCommentCount), num = n())
```

Use group by to get stats by a grouping of the data

```{r}
mean_metric_per_var <- function(metric, variable){
  result = integer(length(variable))
  for (v in variable){
    result[v] <- means(metric[variable == v])
  }
  return(result)
}
```

dplyr

```{r}
#research %>% group_by(journal, year) %>% summarise(tweets_mean = mean(backtweetsCount)) .....
```

```{r}
tweets_per_journal <- research %>%
  group_by(journal) %>%
  summarise(mean_tw = sd(backtweetsCount),
            sd_tw = sd(backtweetsCount),
            num=n(),
            SEM=sd_tw/sqrt(num)) %>%
  select(-sd_tw)

tweets_per_journal
```

