---
title: "plots_ggplot2"
author: "Peter Yang"
date: "1/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

Load data:

```{r}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim="\t")
research <- filter(counts_raw, articleType == "Research Article")
```

# Mapping data to plot aesthetics

Goals:
1. Map columns of a data frame to plot aesthetics using aes
2. Add layers to plot

```{r}
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount, y = wosCountThru2011))
p <- p + geom_point()
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011)) + geom_point()
p
```

Changes aes of geometry

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011)) + geom_point(aes(color = journal))

p
```

Change size of points by variable
```{r}
p <- ggplot(research, aes(x=log(pdfDownloadsCount), y=wosCountThru2011)) + geom_point(aes(size=authorsCount, color = journal))
p
```

Change alpha by variable

```{r}
p <- ggplot(research, aes(x=log(pdfDownloadsCount), y=wosCountThru2011)) + geom_point(aes(size=authorsCount, color = journal, alpha = daysSincePublished))
p
```

Change color on all points not based on data condition

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011)) + geom_point(color="red", aes(size=authorsCount))
p
```

Add another geometry later:

Smoothed line

aes for each specific layer

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011)) + geom_point(aes(color=journal)) + geom_smooth()
p
```

Color both layers by journal

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011, color=journal)) + geom_point() + geom_smooth()

p
```

```{r}
#p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011, color=journal)) + geom_point() + geom_smooth()
p_challenge <- ggplot(research, aes(x=daysSincePublished, y=wosCountThru2011)) + geom_point(alpha = 0.5, aes(color = journal)) + geom_smooth(color = "red")
p_challenge
```

# Control plot scales

Scale how the data maps to the aes
Control axes
Control color scale

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount, y=wosCountThru2011)) + geom_point(aes(color = journal)) + geom_smooth()
p <- p + scale_x_log10() + scale_y_log10()
p

```

```{r}
p <- ggplot(research, aes(x=log10(pdfDownloadsCount + 1), y=log10(wosCountThru2011 + 1))) + geom_point(aes(color = journal)) + geom_smooth()
p
```

```{r}
p <- ggplot(research, aes(x=log10(pdfDownloadsCount + 1), y=log10(wosCountThru2011 + 1))) + 
  geom_point(aes(color = journal, alpha = 0.5)) + geom_smooth() + scale_x_continuous(breaks = c(1,3), labels = c(10,100)) +
  scale_y_continuous(breaks = c(1,3), labels = c(10,100))
p
```

Change the color scale

```{r}
p + scale_color_grey()

```

```{r}
p + scale_color_manual(values = c("red", "yellow", "orange", "purple", "blue", "green", "pink")) 

```


```{r}
p <- ggplot(research, aes(x=sqrt(pdfDownloadsCount), y=sqrt(wosCountThru2011))) + 
  geom_point(aes(color = journal, alpha = 0.5)) + geom_smooth() + scale_color_brewer(palette = "Accent") +
  labs(x="square root of downloads", y = "square root of citation count thru 2011", title = "test title") + 
  theme(plot.title = element_text(hjust = 0.5))
p

```