---
title: "Trial report"
author: "Peter Yang"
date: "1/12/2019"
output: 
  html_document:
    # This sets global options
    toc: true
    number_sections: true
    #theme: cerulean
    #highlight: espresso
    #fig_width: 3
    #fig_height: 3
    fig.caption : yes
    
---

<!-- toc: true means table of contents; include=FALSE means chunk won't show -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!--  message=FALSE means only show code but not messages (results='hide' means no output result) -->
```{r load_library, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Initial Analysis

## Reading in data on plos papers

```{r read_data}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim = "\t")

```

## Diagnostic plots

<!-- customize size of figure just for this chunk -->
```{r histogram_weeks_since_published, fig.height = 3, fig.width = 5, fig.cap = "Figure 1: Weeks since publication per paper"}
weeks_since_pub <- counts_raw$daysSincePublished / 7
hist(weeks_since_pub, xlab = "# of weeks", main = "Weeks since Publication")
```

## Some vague state,ents about data

Here's an example of a math function: $y = x^2$

This daraset consists of `r dim(counts_raw)[1]` publications

<!-- comments comments (this does not show up in document) -->

```{r scatterplot_pubs_vs_citations}
plot(counts_raw$daysSincePublished, counts_raw$wosCountThru2011)
```

## Exercise:
  1. plot a histogram of favorite variable in dataset
  2. make scatter plot of variable against citation counts
    a. comment on whether it has a marked effect on citation counts
    b. if appropriate create a scatter plot of log-transformed values
  3. verify the hitograms look nice
  

```{r histogram_journal}
hist(log(counts_raw$pdfDownloadsCount), main = "log(PDF download count)")
```

```{r scatterplot_wosCountThru2011_vs_pdfDownloadsCount}
plot(log(counts_raw$pdfDownloadsCount), counts_raw$wosCountThru2011, xlab = "log(number of pdf downloads)", ylab = "number of citations since 2011", main = "citations vs log(pdf downlaods)")
```

```{r barplot_trial}
journals <- table(counts_raw$journal)
barplot(journals, xlab = "journal name", ylab = "number of pubs")
```

When a line of code is long, use SHIFT+ENTER for new line
```{r}
library(plyr)
library(ggplot2)

avg_cite_by_journal <- ddply(counts_raw, .(journal), summarize,  citation_avg=mean(wosCountThru2011))
ggplot(avg_cite_by_journal, aes(y=citation_avg, x=journal)) + geom_bar(stat="identity") + 
  geom_text(aes(label=format(citation_avg, digits = 2)), vjust=1.6, color="white", size=3.5) + 
  theme_minimal() + labs(x="Name of Journal", y = "Avg citation per Paper", title = "Test ggplot") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
counts_raw$authorsCount
counts_raw[, "authorsCount"]
counts_raw[["authorsCount"]]

data <- counts_raw$authorsCount
plot(x = counts_raw$wosCountThru2011, y=data)

data_column <- "authorsCount"
log_data <- log(counts_raw[[data_column]])
plot(x=counts_raw$wosCountThru2011, y=log_data)

# Create a new function that does things
plot_log_data <- function(input_dataframe, column){
  data_column <- input_dataframe[[column]]
  log_data_column <- log(data_column)
  return(log_data_column)
}

plot_log_data(input_dataframe = counts_raw, column = "deliciousCount")
```


```{r}
pub_date <- counts_raw$pubDate
head(pub_date)
num_unique <- function(x){
  unique_data <- unique(x)
  length_unique <- length(unique_data)
  print("I'm in num_unique!")
  return(length(unique_data))
}

num_unique(x = counts_raw$pubDate)


```


# Debug test
After calling debug(function), next time when you use "function", it enters debug mode
Use undebug(function) to exit this mode

```{r}
debug(num_unique)
num_unique(x = counts_raw$doi)
```


Use stop command to prevent errors
```{r}
column_num_unique <- function(data,column_name){
  data_column <- data[[column_name]]
  if(is.null(data_column)){
    stop("No such column in data!")
  }
  unique_entries <- num_unique(data_column)
  return(unique_entries)
}
#debug(column_num_unique)
column_num_unique(data = counts_raw,column_name = "doi")
```

### For loops and for fun
seq_along returns index sequence of that length 
x <- c(2 4 6 8 20)
seq_along(x) = 1 2 3 4 5

```{r}
names_of_columns <- colnames(counts_raw)
my_column <- "doi"
column_num_unique(data = counts_raw,
                  column_name = my_column)
unique_col_ct <- function(data){
  data_column_names <- colnames(data)
  output_length <- length(data_column_names)
  result <- integer(output_length)
  for(i in seq_along(data_column_names)){
    my_column <- data_column_names[i]
    result[i] <- column_num_unique(data,my_column)
  }
  print("we finished!")
  return(result)
}
```

```{r}
x <- c(8,9,3,5,7)
x[4]
y <- c(1,2)
x[y]
z <- length(x)
x[z]
```

We can also subset using double bracket syntax to get *one* element back 
```{r}
x[[1]]
#x[[y]]
```

## Indexing dataframes
Dataframes can also be indexed with single bracket '[' and double bracket '[[' syntax depending on what we want

```{r}
counts_raw[,1]
y <- c(1,2)
counts_raw[,y]
counts_raw[,c("doi", "daysSincePublished")]
counts_vector <- counts_raw[["doi"]]
doi_dataframe <- counts_raw[,"doi"]
```



```{r}
counts_columns <- c("facebookShareCount", "facebookLikeCount")
max_of_each_col <- function(data, count_colnames){
  number_of_columns <- length(count_colnames)
  result <- integer(number_of_columns)
  for (i in seq_along(count_colnames)){
    my_column <- count_colnames[i]
    my_vector <- data[[my_column]]
    result[i] <- max(my_vector, na.rm = TRUE) # na.rm=True removes NA values, if not removing, max will give NA as output
  }
  return(result)
}

#debug(max_of_each_col)
max_of_each_col(counts_raw, counts_columns)

```